FraCaS: Temporal Analysis

### 2020-06-04 paper2020/paper.tex
wibble

For the purpose of this paper, the important point is that GF the
syntax aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
unambiguously natural-language text. \cref{ex:onelingex}
shows an example abstract syntax tree and its realisation in English.

@@

\paragraph{Temporal Semantics}
\todo{More general notion of temporal semantics? Should this be moved somewhere else?}

@@

In montagovian semantics, (intransitive) verbs are one-place
predicates; in types, they are functions from entities to propositions
($e \to t$).  Our basic approach is to generalise the interpretation of 
verbs, so that it takes two
additional time parameters, one corresponding with the starting time
of the action and one corresponding to its stopping time ($(time × time × e) \to t$). For example,
if John walked between $t_0$ and $t_1$, we would have:
$walk(john,t_0,t_1)$. From now on we will call an interval of time
points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
using a simple newtonian model of time, corresponding to a layman
intuition.)

@@

We adjust the montagovian semantics so that the interpretation of
every category (propositions, verb phrases, etc.) takes a
\emph{temporal context} as an additional parameter, serves as a time
reference for the interpretation of all time-dependent semantics
within the phrase.  This context propagates through the compositional
interpretation down to lexical items with atomic representation
(verbs). By default, every interpretation passes the temporal context
down to its components without changing it. However some key elements
will act on it on nontrivial ways, which we proceed to detail below.

This temporal context is an \emph{optional} timespan. That is, it can
be an timespan or an explicitly unspecified context.
\newcommand\nospan{\ensuremath{─}}
\newcommand\anyspan{\ensuremath{\_}} The timespan in the context is
optional because, in certain situations, the semantics is different
depending on whether a timespan has been specified externally or not,
as we explain below. A non-present timespan will be represented as
\nospan{}. If a semantic function does not depend on the temporal
context at all, we will write \anyspan{} instead.

@@

\onelingex{A scandinavian won the nobel prize.\label{ex:onelingex}
\\
{\small\begin{multline*}
\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
\end{multline*}
}}
In our semantics we deal only with present and simple past
tenses. \jp{Is this enough for theory?} Indeed we find that FraCas
does not exercise additional specific tenses (if a more complicated
tense is used, the additional information is also carried by adverbs
or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.

>>>>

For the purpose of this paper, the important point is that GF the
syntax aims at a balance of sufficient abstraction to provide a
semantically-relevant structure, but at the same time it embeds
sufficiently many syntactic features to be able to reconstruct
unambiguously natural-language text. \cref{ex:nobel}
shows an example abstract syntax tree and its realisation in English.

@@

\paragraph{Temporal Semantics}
\jp{More general intro to temporal semantics? Should this be moved
  somewhere else?}

@@

In montagovian semantics, (intransitive) verbs are one-place
predicates; in types, they are functions from entities to propositions
($e \to t$).  Our basic approach is to generalise the interpretation of 
verbs, so that it takes two
additional time parameters, one corresponding with the starting time
of the action and one corresponding to its stopping time ($(e × time × time) \to t$). For example,
if John walked between $t_0$ and $t_1$, we would have:
$walk(john,t_0,t_1)$. From now on we will call an interval of time
points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
property $t_0 ≤ t_1$ (it starts no later than it stops). (We are thus
using a simple newtonian model of time, corresponding to a layman
intuition of a linear constant flow of time.)

@@

We adjust the montagovian semantics so that the interpretation of
every category (propositions, verb phrases, etc.) takes a
\emph{temporal context} as an additional parameter, which serves as a time
reference for the interpretation of all time-dependent semantics
within the phrase.  This context propagates through the compositional
interpretation down to lexical items with atomic representation
(verbs). By default, every interpretation passes the temporal context
down to its components without changing it. However some key elements
will act on it on nontrivial ways, which we proceed to detail below.

This temporal context is an \emph{optional} timespan. That is, it can
be an timespan or an explicitly unspecified context.
\newcommand\nospan{\ensuremath{-}}
\newcommand\anyspan{\ensuremath{\_}}\todo{better notation} The timespan in the context is
optional because, in certain situations, the semantics is different
depending on whether a timespan has been specified externally or not,
as we explain below. A non-present timespan will be represented as
\nospan{}. If a semantic function does not depend on the temporal
context at all, we will write \anyspan{} instead.

@@

\onelingex{A scandinavian won the nobel prize.\label{ex:nobel}
\\
{\small\begin{multline*}
\varid{useCl}\, \varid{past}\, \varid{pPos}\\ (\varid{predVP}\, (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{scandinavian\_CN})\\ (\varid{complSlash}\, (\varid{slashV2a}\, \varid{win\_V2})\\ (\varid{detCN}\, (\varid{detQuant}\, \varid{indefArt}\, \varid{numSg})\\ \varid{nobel\_prize\_CN})))
\end{multline*}
}}
In our semantics we deal only with present and simple past
tenses. \jp{Is this enough for theory?} Indeed we find that FraCas
does not exercise additional specific tenses (if a more complicated
tense is used, the additional information is also carried by adverbs
or adverbial phrases, in a more specific way.) It is not difficult to see how other tenses can be captured under the same general framework.
